---
title: "Bumblebee Lab 14"
author: "Rachael Robinson, Manasi Raol, Brady Kiesling, William Brickowski"
date: "4/24/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Lab 14 (Lab 4 Redo)

## Team Questions:
  * Does smoking lead to babies born prematurely?
  * Does smoking lead to low birth weight? 

## Why are these questions important?
Reason: These questions are important because health behaviors during pregnancy are often a topic of interest. Determining what behaviors are more or less likely to result in healthy babies is important because it can inform women and families to improve health outcomes. Smoking in particular is an important question of interest due to the number of smokers during the 1960s, in which this data was collected.
  * This reason is important because it has wide-reaching impacts on generational health. Widespread habits can result in health affects on entire populations of babies, which could lead to consequences as this population of babies grows into adulthood. Trends in low birth weights or premature births could eventually impact the quality of life, labor, and health of larger populations.


```{r include=FALSE}
library(tidyverse)
babies <- read_csv("https://raw.githubusercontent.com/ervance1/Sp2018-Data-Science-repo/master/babies2a.dat")
babies <- rename(babies, bwtoz = `bwt/oz`) 
babies = na.omit(babies)
```


## Preliminary Team Questions:
  * Which variables are similar between smokers and nonsmokers? Which are different?

## Team Plots for Prelimary Questions:

```{r}
corr_test <- function(value1, value2, n){
  test1 <- sample(value1, n, replace = TRUE)
  test2 <- sample(value2, n, replace = TRUE)
  relate <- cor(test1,test2)
  return(relate)
}

corr_parity <- corr_test(smoke$parity,nosmoke$parity,n=200)
corr_mrace <- corr_test(smoke$mrace,nosmoke$mrace,n=200)
corr_mage <- corr_test(smoke$mage,nosmoke$mage,n=200)
corr_med <- corr_test(smoke$med,nosmoke$med,n=200)
corr_mht <- corr_test(smoke$mht,nosmoke$mht,n=200)
corr_mpregwt <- corr_test(smoke$mpregwt,nosmoke$mpregwt,n=200)
corr_drace <- corr_test(smoke$drace,nosmoke$drace,n=200)
corr_dage <- corr_test(smoke$dage,nosmoke$dage,n=200)
corr_ded <- corr_test(smoke$ded,nosmoke$ded,n=200)
corr_dht <- corr_test(smoke$dht,nosmoke$dht,n=200)
corr_dwt <- corr_test(smoke$dwt,nosmoke$dwt,n=200)
corr_marital <- corr_test(smoke$marital,nosmoke$marital,n=200)
corr_inc <- corr_test(smoke$inc,nosmoke$inc,n=200)

correlations <- c(corr_parity,corr_mrace,corr_mage,corr_med,corr_mht,corr_mpregwt,corr_drace,corr_dage,corr_ded,corr_dht, corr_dwt,corr_marital,corr_inc)

babyname <- babies %>%
  select(parity, mrace, mage, med,mht,mpregwt,drace,dage,ded,dht,dwt,marital,inc)

varnames <- colnames(babyname)

ggplot() +
  geom_point(mapping = aes(x=varnames,y=correlations)) +
  labs(title = "Variable correlation between smokers and nonsmokers", x = "Variable")
```

```{r}
babies$smoke <- as.factor(babies$smoke)

map_plot <- function(Y_VAR)
{
    ggplot(data=babies,aes(x=smoke,y=Y_VAR,fill=smoke)) +
    geom_bar(aes(y=..count../sum(..count..),fill=smoke),position="dodge")+
    xlab('Smoke: 0 "never" and 1"smokes now"')+
    ylab('Normalized Count')+
    ggtitle(paste(Y_VAR,'by smokers and non-smokers', sep = " ", collapse = NULL))
}

vector=c('gestation','bwtoz','parity','mage','med','mht','mpregwt','dage','ded','dht','dwt','marital','inc','time','Premature')

map(vector,map_plot)
```

## Conclusions for Prelimary Questions:

Looking at the areas where the correlations are strong between the two variables, we can see the differences and similarities between smokers and nonsmokers in the dataset. Overall, most of the correlations are low, meaning we have low prediction power in between groups. For instance, knowing about the mother ages of the smokers does little to tell us about the ages of the nonsmokers, since they are nearly uncorrelated. Thus, high positive correlation means the variables are similar in behavior, and high negative correlation means the variables are similar in behavior but in opposite directions. From these generalizations, we can conclude from this plot that in our dataset, the race of the dad, the age of the mother, the marital status of the mother, are all very different in the dataset between the smokers and the nonsmokers. The rest of the variables are either non-generalizable or similar in both groups.


## Question 1

```{r}
babies_smoke = babies %>% filter(smoke == 1)
babies_notSmoke = babies %>% filter(smoke == 0)

lmod_smoke = lm(gestation ~ bwtoz+parity+mage+med+mht+mpregwt+dage+ded+dht+dwt+marital+inc+time+Premature, data = babies_smoke )

lmod_notSmoke = lm(gestation ~ bwtoz+parity+mage+med+mht+mpregwt+dage+ded+dht+dwt+marital+inc+time+Premature, data = babies_notSmoke)

summary(lmod_smoke)
summary(lmod_notSmoke)
```



## Question 2


```{r}
# perm function

perm_mean <- function(perms = 1000, values, n1)
{
  ## Variables ##
  # perms: The number of permutations 
  # values (num): 
  # n1 (int): Size of group 1
  ###############
  
  # Step 1:
  # Create vector of zeroes of length "perms" to store
  # permuted mean differnces
  
  mean_diff <- vector("double", length = perms)
  
  # Loop throught number of permutations
  for (i in c(1:perms))
  {
    # Step 2:
    # Randomly separate vector "values" into disjoint 
    # groups of size "n1" and "length(values) - n1" respectively
    
    mean_sep1 <- sample(values, n1, replace = T)
    mean_sep2 <- setdiff(values, mean_sep1)
    
    
    # Step 3:
    # Compute the sample means for the two groups from 
    # step 2
    
    mean_value1 <- mean(mean_sep1)
    mean_value2 <- mean(mean_sep2)
    
    # Step 4: 
    # Compute the difference in sample means, store the
    # value in the vector from step 1
    
    mean_diff[i] <- mean_value1 - mean_value2
  }
  # Step 5:
  # Return new updated vector, created in step 1
    
  return(mean_diff)
  
}


smoke_test <- babies %>%
  filter(smoke == 1) %>%
  filter(Premature == 0) %>%
  select(bwtoz)


nosmoke_test <- babies %>%
  filter(smoke == 0) %>%
  filter(Premature == 0) %>%
  select(bwtoz)


permute <- perm_mean(perms = 1000, babies$bwtoz, length(nosmoke_test))

mean_smoke <- mean(smoke_test$bwtoz)
mean_nosmoke <- mean(nosmoke_test$bwtoz)

meandifference <- mean_nosmoke - mean_smoke


ggplot() +
  geom_histogram(mapping = aes(x=permute)) +
  geom_vline(mapping = aes(xintercept = meandifference, color = "Original sample mean difference")) +
  labs(title = "Two mean permutation test for birth weight difference in smoking vs non-smoking mothers", subtitle = "1000 permutation test, red line represents true difference in mean birth weight", x = "Mean difference")
```

Percentile for question 2:

```{r}
percentile.data <- sum(permute<meandifference)/1000

print(percentile.data)
```

## Conclusion

Question 1 - According to regression analysis among the following variables as the predictors bwtoz+parity+mage+med+mht+mpregwt+dage+ded+dht+dwt+marital+inc+time+Premature and gestation as the response, sepeated by smokers and non-smokers, the intercept term suggests that gestation is longer for non-smokers. This shows that more babies are likely to be born premature when that babies mother smokes. Additionally, the regression analysis shows what predictors contribution to longer or shorter periods of gestation for babies coming from a mother who smokes or a mother who does not. The significant predictors of gestation period for smokers are bwtoz and mht. If mht increases by 1 unit on average for a smoker, gestation period is reduced by 1.04 units. If bwtoz increases by 1 unit on average for a smoker, gestation period increases by 0.24 units. For non-smokers, the significant predictors are bwtoz, parity, and marital. For non-smokers, if marital is 1 then gestation increases on average by 6.66 units. If parity increases by 1 unit on average, then gestation is reduced by 1.63 units. If btwoz is increased by 1 unit on average then gestation is increased by 0.11 units.

Question 2 - According to our permutation test, we did not see a significant difference in the true means for the birthweights of smoking versus non-smoking mothers. On completion of the two-mean permutation test and plotting, we saw that our test statistic (difference in mean birth weights) fell almost directly in the middle of the distribution. This suggests that statistically, there is no significant difference in this dataset between smoking and nonsmoking mothers.

## Recommendations:

### Recommendation for Q1
Because many of the predictors were not significant, I believe further multi-variable regression models should be built by removing all predictors at a particular significance level in order to more accurately depict the relationship between the respective predictors (for both smokers and non-smokers) and gestation period. An F test can be used in to compare full linear models and reduced ones to determine which model is better.

### Recommendation for Q2

Because our testing showed a non-significant difference in birth weight due to smoking, we recommend that data collection should be furthered in order to gain more insight into this issue. Furthermore, we suggest that additional statistical tests be run with control for other variables in order to tease out if there is a significant difference being confounded by another variable in the dataset. On the question of birthweight, at this moment we cannot make a specific recommendation on behavior changes based on this data.

## New techniques used/comparison to first answer

Question 1 - In order to answer this question, we originally concluded that there is no significant relationship between whether the mother smokes and premature births by looking at the boxplots, the distibution is roughly the same, even though the median gestational time for smokers is lower, the values for the first and 3rd quartiles are roughly the same. However, in order to answer this question with newer techniques, I performed regression for smokers and non-smokers, respectively, with gestation as the response. I then looked at the intercept term of the regerssion summary which represents the average gestation period in our data. While the intercept term was slightly higher for non-smokers, it was not significantly different. Hence, by regression analysis, there is no real difference in gestation period between smokers and non-smokers. This affirms our original conclusion on the question.

Question 2 - On our original lab we created a scatterplot to look at birthweights across different gestational ages to determine whether smoking had an impact. When controlling for gestational age, we saw a weak relationship between smoking and birth weight. In this lab, instead of plotting and making conclusions based on visualisations, we employed a statistical test to try to identify a difference in birthweights. Still controlling for premature birth, we saw zero statistical significance to suggest a difference in the two groups. This technique is much more precise, and allowed us to come to a stronger conclusion than the first time around. In comparison to our first answer, we concluded that there is no clear statistical relationship between birthweight and smoking mothers.

## Individual Plots

### William

```{r}
babies$smoke <- as.factor(babies$smoke)
  
  ggplot(data = babies, aes(x = bwtoz, y = gestation, color = smoke)) +  
  geom_point() +                                                     
  ggtitle('Length of gestation in days verses the pre-pregnancy weight of the mother') +                         
  xlab('Pre-pregnancy weight') +                                              
  ylab('Gestation') +                                              
  scale_color_discrete(name = 'Smoke:',                              
                       labels = c('Never Smokes', 'Smokes Now')) +
  geom_smooth()
```
This is a scatter plot whose x variable is the gestation period in days and whose y value is the bwtoz. The graph shows data for mothers who have never smoked and those who smoke. For both smokers and non-smokers, their is a quite clear linear trend in which a larger bwtoz corresponds to a larger gestation period. This relationship makes sense because the longer a baby is in the womb the larger it can get. Using regression analysis to drive my further research and graphing was a great way to do it for it shows the relationship between many predictors and gestation for our respective groups.

### Brady

### Manasi
```{r}
library(tidyverse)
babies <- read_csv("https://raw.githubusercontent.com/ervance1/Sp2018-Data-Science-repo/master/babies2a.dat")
babies <- rename(babies, bwtoz = `bwt/oz`)
babies <- na.omit(babies) #removes NAs from list of values 


ggplot(data = babies, aes(x = bwtoz, fill = factor(Premature))) +
  geom_dotplot(binwidth = 0.1) +
  ggtitle("Premature vs Non-premature babies For Specific Birth weights")
```

### Rachael

Subquestions: What is the relationship between education level and smoking habits? Is there a significant relationship between birthweights and the mother's education level?

Why is this question important: This question is important because it extends from the information we learned in the last lab. As a team, we determined that in this data, there isn't a relationship between smoking and birthweight. So, this question attempts to understand why that may be the case. This relates to the team question because it is good to definitively conclude what is not related in the data, but even more insightful to then say what is related in the data. To answer this question, I used our new programming tools of writing functions and employing them to answer statistical questions. I also used new functions to calculate 

Below is the correlation between mother's smoking habits and education level. This indicates a weakly negative relationship between the two, and thus that the more the mother smokes, the less her education level is likely to be. This correlation isn't significant on it's own, so further testing would need to be done to show more significance. From this figure, I can conclude a casual relationship between the two, but possibly not significant enough to confound results.

```{r}
babyed <- babies %>%
  filter(smoke == 0 | smoke == 1) %>%
  filter(med != 9, med != 7)
cor(babyed$smoke, babyed$med)
```

Below is the correlation between mother's education level and birthweight. This will be tested against a 1000 permutation correlation plot and percentile.

```{r}
realcor <- cor(babyed$med, babyed$bwtoz)
realcor
```

```{r}
perm_cor <- function(perms = 1000, x, y)
{
  ## Variables ##
  # perms: The number of permutations 
  # x: Vector 1 - for computing correlation
  # y: Vector 2 - for computing correlation
  ###############
  
  # Step 1:
  # Create vector of zeroes of length "perms" to store
  # permuted mean differnces
  
  corr_result <- vector("double", length = perms)
  
  # Loop throught number of permutations
  for (i in c(1:perms))
  {
    # Step 2:
    # Randomly mix up the values in the vector "y"
    
    y <- sample(y)
    
    # Step 3:
    # Compute the correlation between x and the randomly mixed
    # up y-vector. Store this value in the vector from step 1.
    
    corr_result[i] <- cor(x,y)
  }
  
  # Step 4:
  # Return new updated vector, created in step 1
  return(corr_result)
}



permuted <- perm_cor(perms = 1000, babyed$med, babyed$bwtoz)

ggplot() +
  geom_histogram(mapping = aes(x=permuted)) +
  geom_vline(mapping = aes(xintercept = realcor, color = "Original sample correlation")) +
  labs(title = "Correlation permutation test for birth weight difference and education levels", subtitle = "1000 correlation test, red line represents true correlation between birth weight and education", x = "Correlation")

```



Percentile calculation: 

```{r}
percentite <- sum(permuted<realcor)/1000

print(percentite)
```

From the histogram and subsequent percentile calculation, it's clear that we can likely reject the null and assume that the correlation between birthweight and education level. The true correlation we calculated was in the tail of the histogram, suggesting there is a correlation, and there is a weakly positive relationship between education level and birthweight. With smoking as a possible confounder, this is an interesting and significant result.

Overall, we reject the null and assume that the correlation between education and birthweight is nonzero.





### Who did what

### William

Within this lab, I generated the preliminary graphs. I had originally written seperate code for each of the box plots generated, however, in this lab, I imlemented this using new techniques that we had learned. I wrote a generic function to create a boxplot with smoking as the x variable and an arbitraty variable for the y variable. I then applied the map function to a vector of all the possible y variables which produced all of the same plots in many less lines of code. I also performed regression analysis for smokers and non-smokers, respectively, using gestation as the response and various other variables as the predictors. I used this analysis to answer our first subquestion.I also graphed a scatter plot relating bwtoz and gestation period for both smokers and non-smokers because this was a significant predictor in the regression analysis for both of the groups - smokers and non-smokers. The regression analysis can be used to better understand what causes variations in gestation period for the two respective groups. Bwtoz clearly is linearly related to gestation period. There are also other predictors which are uniquely predictive for smokers or non-smokers.

### Brady

### Manasi
For this lab I contributed to question two by making a plot to show the ratio of premature verus non premature babies in a range of different birth weights.

### Rachael


For this lab, I developed my own subquestion and answered it with new techniques. I did some analysis and created a plot to demonstrate my findings. Additionally, for this lab I developed the preliminary analysis and subsequent discussion, the team plot for question 2, and following analysis, conclusions, and recommendations.