---
title: "bumblebeeLab10"
author: "Manasi Raol"
date: "3/21/2019"
output: html_document
---

#Lab 10

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
answers <- read_csv("Answers_trunc.csv")
questions <- read_csv("Questions_trunc.csv")
```

##Team Part
What features affect the score the most?

#Team Plots

##Individual Findings

###William

###Manasi

```{r}
library(tidyverse)
answers <- read_csv("Answers_trunc.csv")
questions <- read_csv("Questions_trunc.csv")

#feature 1 (answers) 

no <- sum(str_count(answers$Body,pattern = "No"))
yes <- sum(str_count(answers$Body,pattern = "Yes"))
ggplot()

#feature 1 (questions)   
python <- sum(str_count(questions$Body,pattern = "Python"))
sql <- sum(str_count(questions$Body,pattern = "SQL"))

answers %>%
  mutate(no)

answers %>%
  mutate(yes)

ggplot(data = answers, aes(x=no, y= Score)) +
  geom_line()

ggplot(data = answers, aes(x=no, y= Score)) +
  geom_line()

questions %>%
  mutate(python)

questions %>%
  mutate(sql)

ggplot(data = questions, aes(x=python, y= Score)) +
  geom_line()

ggplot(data = answers, aes(x=sql, y= Score)) +
  geom_line()


```

With my features, I wanted to see wheteher more people had questions about Python or SQL so I added one to my Python and SQL variables everytime the word Python or SQL was in the the question. I did this by using str_count. For answers, I wanted to compare whether people got answers they wanted by counting how many times No or Yes was included in the answers. Overall there were more questions about Python and more answers with the word "No" in them.

###Brady

###Rachael


## Who did what

###William

###Manasi

###Brady

###Rachael



