---
title: "Lab4_Bumblebee"
author: "Rachael Robinson, William Brickowski, Manasi Raol, Brady Kiesling"
date: "2/8/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab 4

## Team Questions:
  * Does smoking lead to babies born prematurely?
  * Does smoking lead to low birth weight? 

## Why are these questions important?
Reason: These questions are important because health behaviors during pregnancy are often a topic of interest. Determining what behaviors are more or less likely to result in healthy babies is important because it can inform women and families to improve health outcomes. Smoking in particular is an important question of interest due to the number of smokers during the 1960s, in which this data was collected.
  * This reason is important because it has wide-reaching impacts on generational health. Widespread habits can result in health affects on entire populations of babies, which could lead to consequences as this population of babies grows into adulthood. Trends in low birth weights or premature births could eventually impact the quality of life, labor, and health of larger populations.

## Team Plots:


### Question 1
```{r}
ggplot(data=babies)+
  geom_boxplot(aes(y=gestation, x=as.factor(smoke)),outlier.shape = NA)+
  geom_hline(yintercept=270)+
  ylim(250,320)+
  labs(x="Non-smoking vs smoking", y="days of gestation", title="Distibution of gestation days")
```  
The boxplot represents days of gestation, and the line across represents 270 days. Babies born under that line are considered premature. Even though the median birth day for smokers is slightly lower, the lengh of the line under the 270 mark is roughly the same for both smoking mothers and non-smoking mothers. This means that there is no significant difference in the proportion of smoking and non-smoking mothers giving birth prematurely. This plot seems to refute the 1989 claim that mothers smoking causes premature births.
  
### Question 2

```{r}
ggplot(data = babies) +
  geom_jitter(outlier.shape = NA, mapping = aes(x=gestation, y=bwtoz, color = as.factor(smoke)), alpha = .4) +
  geom_smooth(na.rm = T, method = "lm", se = F, mapping = aes(x=gestation, y=bwtoz, color = as.factor(smoke))) +
   xlim(270,350) +
  theme_classic() +
  xlab("Gestation (days)") +
  ylab("Birth weight (oz)") +
  ggtitle("Observation of smoking and it's impact on birth weight across gestational ages") +
  scale_colour_discrete(name = "Smoking status", breaks = c("0", "1"), labels = c("Non-smoking", "Smoking"))

ggplot(data = babies) +
  geom_jitter(outlier.shape = NA, mapping = aes(x=gestation, y=bwtoz, color = as.factor(smoke)), alpha = .4) +
  geom_smooth(na.rm = T, method = "lm", se = F, mapping = aes(x=gestation, y=bwtoz, color = as.factor(smoke))) +
   xlim(100,350) +
  theme_classic() +
  xlab("Gestation (days)") +
  ylab("Birth weight (oz)") +
  ggtitle("Observation of smoking and it's impact on birth weight across gestational ages") +
  scale_colour_discrete(name = "Smoking status", breaks = c("0", "1"), labels = c("Non-smoking", "Smoking"))
```
This first plot is a scatterplot showing the birth weights across different gestational ages in our data. Then, a linear model was fitted to the points of the nonsmokers and the smokers. In question 1, we determined no significant relationship between smoking and premature birth, and a premature baby is much more likely to have a lower than average birth weight. To control for the effect that a premature birth would have on gestational birth weight, the gestational scale was set at a minimum of 270 days, the cutoff determining a mature baby. Both points follow a weakly linear pattern. The second plot represents the entire range of gestational values, and again shows a weakly linear relationship.

## Conclusions:

  * Answer to first question
  * Based on our graph and our conclusion to the first question, we conclude that there is a minimal relationship between smoking and birth weights. Our first conclusion states that there is no significant relationship between smoking and premature births. When controlling for premature birth, our plot shows that the non-smoking mothers had a slightly higher average of birth weight across mature gestational ages. However, these relations are weakly linear, and the signficance of the difference is only truly present from gestational ages 270 to 290. Looking at the second plot, where premature birth was not controlled for, the relationship seems much stronger between smoking and birth weight across gestational ages. However, several outliers are visible in the non-smoking group, making the control for premature birth a valid control. Ultimately, due to the ambiguity in premature birth and the outliers present in the data, we find a minimal relationship between smoking and birth weight.
  
## Recommendations:

### Recommendation for Q1

### Recommendation for Q2

Given our data, we recommend that women generally avoid smoking while pregnant, as we identified a weakly linear relationship between smoking and birth weight. We also recommend that the relationship between gestation and birth weight is further investigated in order to account for a possibly nonlinear relationship. Based on our data, we recommend that the Surgeon General adapts their 1989 report to reflect that the data only supports the affect of smoking at gestational ages considered mature (>270).

## Preliminary Team Questions:
  * Which variables are similar between smokers and nonsmokers? Which are different?

## Team Plots for Prelimary Questions:
```{r}
library(tidyverse)
babies <- read_csv("https://raw.githubusercontent.com/ervance1/Sp2018-Data-Science-repo/master/babies2a.dat")
babies <- rename(babies, bwtoz = `bwt/oz`) 
babies = na.omit(babies)
babies$smoke <- as.factor(babies$smoke)

ggplot(data=babies,aes(x=smoke,y=gestation,fill=smoke)) +
  geom_bar(aes(y=..count../sum(..count..),fill=smoke),position="dodge")+
  xlab('Smoke: 0 "never" and 1"smokes now"')+
  ylab('Normalized Count')+
  ggtitle('Length of gestation by smokers and non-smokers')

ggplot(data=babies,aes(x=smoke,y=bwtoz,fill=smoke)) +
  geom_bar(aes(y=..count../sum(..count..),fill=smoke),position="dodge")+
  xlab('Smoke: 0 "never" and 1"smokes now"')+
  ylab('Normalized Count')+
  ggtitle('Birth weight of baby by smokers and non-smokers')

ggplot(data=babies,aes(x=smoke,y=parity,fill=smoke)) +
  geom_bar(aes(y=..count../sum(..count..),fill=smoke),position="dodge")+
  xlab('Smoke: 0 "never" and 1"smokes now"')+
  ylab('Normalized Count')+
  ggtitle('Total number of previous pregnancies by smokers and non-smokers')

ggplot(data=babies,aes(x=smoke,y=mage,fill=smoke)) +
  geom_bar(aes(y=..count../sum(..count..),fill=smoke),position="dodge")+
  xlab('Smoke: 0 "never" and 1"smokes now"')+
  ylab('Normalized Count')+
  ggtitle('Mother age by smokers and non-smokers')

ggplot(data=babies,aes(x=smoke,y=med,fill=smoke)) +
  geom_bar(aes(y=..count../sum(..count..),fill=smoke),position="dodge")+
  xlab('Smoke: 0 "never" and 1"smokes now"')+
  ylab('Normalized Count')+
  ggtitle('Mother educaiton by smokers and non-smokers')

ggplot(data=babies,aes(x=smoke,y=mht,fill=smoke)) +
  geom_bar(aes(y=..count../sum(..count..),fill=smoke),position="dodge")+
  xlab('Smoke: 0 "never" and 1"smokes now"')+
  ylab('Normalized Count')+
  ggtitle('Mother height by smokers and non-smokers')

ggplot(data=babies,aes(x=smoke,y=mpregwt,fill=smoke)) +
  geom_bar(aes(y=..count../sum(..count..),fill=smoke),position="dodge")+
  xlab('Smoke: 0 "never" and 1"smokes now"')+
  ylab('Normalized Count')+
  ggtitle('Mother weight by smokers and non-smokers')

ggplot(data=babies,aes(x=smoke,y=dage,fill=smoke)) +
  geom_bar(aes(y=..count../sum(..count..),fill=smoke),position="dodge")+
  xlab('Smoke: 0 "never" and 1"smokes now"')+
  ylab('Normalized Count')+
  ggtitle('Father age by smokers and non-smokers')

ggplot(data=babies,aes(x=smoke,y=ded,fill=smoke)) +
  geom_bar(aes(y=..count../sum(..count..),fill=smoke),position="dodge")+
  xlab('Smoke: 0 "never" and 1"smokes now"')+
  ylab('Normalized Count')+
  ggtitle('Father education by smokers and non-smokers')

ggplot(data=babies,aes(x=smoke,y=dht,fill=smoke)) +
  geom_bar(aes(y=..count../sum(..count..),fill=smoke),position="dodge")+
  xlab('Smoke: 0 "never" and 1"smokes now"')+
  ylab('Normalized Count')+
  ggtitle('Father height by smokers and non-smokers')

ggplot(data=babies,aes(x=smoke,y=dwt,fill=smoke)) +
  geom_bar(aes(y=..count../sum(..count..),fill=smoke),position="dodge")+
  xlab('Smoke: 0 "never" and 1"smokes now"')+
  ylab('Normalized Count')+
  ggtitle('Father weight by smokers and non-smokers')

ggplot(data=babies,aes(x=smoke,y=marital,fill=smoke)) +
  geom_bar(aes(y=..count../sum(..count..),fill=smoke),position="dodge")+
  xlab('Smoke: 0 "never" and 1"smokes now"')+
  ylab('Normalized Count')+
  ggtitle('Marital status of mother by smokers and non-smokers')

ggplot(data=babies,aes(x=smoke,y=inc,fill=smoke)) +
  geom_bar(aes(y=..count../sum(..count..),fill=smoke),position="dodge")+
  xlab('Smoke: 0 "never" and 1"smokes now"')+
  ylab('Normalized Count')+
  ggtitle('Family yearly income by smokers and non-smokers')

ggplot(data=babies,aes(x=smoke,y=time,fill=smoke)) +
  geom_bar(aes(y=..count../sum(..count..),fill=smoke),position="dodge")+
  xlab('Smoke: 0 "never" and 1"smokes now"')+
  ylab('Normalized Count')+
  ggtitle('Time since quiting by smokers and non-smokers')

ggplot(data=babies,aes(x=smoke,y=Premature,fill=smoke)) +
  geom_bar(aes(y=..count../sum(..count..),fill=smoke),position="dodge")+
  xlab('Smoke: 0 "never" and 1"smokes now"')+
  ylab('Normalized Count')+
  ggtitle('Percent of premature child birth by smokers and non-smokers')
```

## Conclusions for Prelimary Questions:
Based on these prelimnary graphs which exam the difference among the majority of the variables in the dataframe between smokers and none smokers, there have been several differences identified. For mothers who do not smoke, their length of gestation is longer. The birth weight of babies born to a mother who does not smoke is larger than for babies born to a mother who does. Mothers who do not smoke tend to be younger and more education, the same goes for the father. Additionally, for both mothers and fathers, on average, those who do not smoke are taller and heavier.


## Individual Plots

### William
```{r}
library(tidyverse)
babies <- read_csv("https://raw.githubusercontent.com/ervance1/Sp2018-Data-Science-repo/master/babies2a.dat")
babies <- rename(babies, bwtoz = `bwt/oz`) 
babies = na.omit(babies)
babies$smoke <- as.factor(babies$smoke)
  
  ggplot(data = babies, aes(x = mpregwt, y = gestation, color = smoke)) +  
  geom_point() +                                                     
  ggtitle('Length of gestation in days verses the pre-pregnancy weight of the mother') +                         
  xlab('Pre-pregnancy weight') +                                              
  ylab('Gestation') +                                              
  scale_color_discrete(name = 'Smoke:',                              
                       labels = c('Never Smokes', 'Smokes Now')) +
  geom_smooth()
```
This is a scatter plot whose x variable is the gestation period in days and whose y value is the pre-pregnancy weight of the  mother. The graph shows data for mothers who have never smoked and those who smoke. For both smokers and non-smokers, their is a quite clear linear trend in which mothers with a larger pre-preganancy weight have a longer gestation period. This is true up until a certain weight, and at this point the trend becomes the opposite. This could describe the part of the sample that is overweight and unhealthy. The distinction between smokers and non-smokers is that the line of best fit for smokers appears to be the same from non-smokers, however, it is shifted down slightly. This means that for the same pre-pregnancy weight, non-smokers have longer gestation periods. Additionally, based on the preliminary graphs which show that mothers who smoke tend to weigh less, this could explain why smokers in general have shorter gestation periods. Again, due to the presence of confounding variables, correlation cannot be confused with causation and further testing would be required to generate a more definite conclusion.

### Brady

### Manasi
```{r}
library(tidyverse)
babies <- read_csv("https://raw.githubusercontent.com/ervance1/Sp2018-Data-Science-repo/master/babies2a.dat")
babies <- rename(babies, bwtoz = `bwt/oz`)
babies <- na.omit(babies) #removes NAs from list of values 


ggplot(data = babies, aes(x = inc, fill = factor(Premature))) +
  geom_dotplot(binwidth = 0.1) +
  ggtitle("Premature vs Non-premature babies For Each Level of Income")
```
This is a dotplot that shows the ratio of premature verus non premature babies in every income level. Overall the middle income levels show about the same ratio. However the lowest income level shows about half of its babies as premature and the two highest income level shows almost no babies as premature.

### Rachael

```{r}
library(tidyverse)
babies <- read_csv("https://raw.githubusercontent.com/ervance1/Sp2018-Data-Science-repo/master/babies2a.dat")
babies <- rename(babies, bwtoz = `bwt/oz`) #renaming `bwt/oz` because the "/" sign requires special care
no_unknown_babies <- subset(babies, med != 9)
ggplot(data = no_unknown_babies) +
  geom_boxplot(mapping = aes(x=factor(med), y=bwtoz, group = med, fill = factor(med)), outlier.shape = NA) +
  geom_hline(yintercept = mean(babies$bwtoz), linetype = "dashed", color = "red", size = 1) +
  theme_minimal()+
  xlab("Education status of mother") +
  ylab("Birth weight (oz)") +
  ggtitle("Education status of mother versus birth weight") +
  scale_fill_discrete(name="Education status",breaks=c("0", "1", "2", "3", "4", "5", "7"),
                         labels=c("less than 8th grade", "8th to 12th grade. did not graduate high school", "high school graduate, no other schooling", "high school graduate + trade school", "high school graduate + some college", "college graduate", "trade school"))
```
This is a boxplot showing the spread of birth weights under the lens of the education status of the mother. As seen in the boxplot, there isn't a glaringly obvious difference between the birth weights given the varying education levels. The red line on the graph indicates the average birth weight across all the data, and most of the education levels are close to this average. However, it is slightly noticeable that some of the lower education levels (in quantity of years) show a lower than average median birth weights. Overall, I can reasonably report that there is no significant relationship present in this data between the mothers education level and birthweights. An extension of this finding is that education and it's relationship to smoking should not confound the relationship between smoking and birth weight.

## Individual contribution and non-defaults for plot generation

### William
Within this lab, I generated the preliminary graphs as well as wrote the conclusion that corresponds to them. We say several interesting relationships between whether the mother smokes with respect to other variables. The one that sparked my individual exploration was that mothers with a smaller pre-pregnancy weight tended to be smokers and mothers who did not smoke weighed more. Another graph that further led to my individual research was one that showed that smokers had shorted gestation periods and non-smokers had longer ones. From these two graphs, I was interest in perhaps determining if weight was a confounding variable. For this reason, I graphed a scatter plot relating pre-pregnancy weight and gestation period for both smokers and non-smokers. The results confirmed my intuition, mothers who weighed more, up until a certain point, had longer gestation periods. With regards to the non-defaults for plot generation, on the plots that I generated within the preliminary section, I normlaized the data through a parameter within the geom_bar function. This had to be done because there were not an equal number of mothers who smoked and who did not smoke. I also added x labels and y labels to these preliminary plots as well as a title. Furthermore, I used the position parameter with the geom_bar function to dodge the bar charts so that smokers and non-smokers could be compared side by side. With respect to my individual plot titled "Length of gestation in days verses the pre-pregnancy weight of the mother", I added the color parameter to the ggplot function so that I could get a scatter plot seperated by color based on smokers and non-smokers. I then added a title and x/y labels. I added a scale title and changed the label names. Additionally, I made use of the geom_smooth function so that lines up best fit could easily be compared to make conclusion upon the data.

### Brady

### Manasi
In this lab, I helped create the prelilmnary graphs to find any trends they may have produced. Additionally I contributed my own indivudal plot as well as findings.

### Rachael

For this lab, I created the team plot for Question 2, as long as assisting with the analysis of Question 1 and the preliminary data. I also wrote the conclusion and recommendation for team question 2. For the report, I assembled and organized the notebook, and wrote about the importance of these research questions. In addition, I created my own individual plot and reported my findings from that plot. For that plot, I restructured the data frame to get rid of extraneous data. I then mapped the education of the mother to a discrete variable in order to make a boxplot. I colored the boxplots based on the mother's education level, and eliminated the outliers to help the presentation of my boxplots. I then created a line to represent the mean birth weight, and manually changed the aesthetics to change the line style, color, and size. Additionally, I used themes, scales, and labels to represent my data clearly.