---
title: "Lab5_Bumblebee"
author: "William Brickowski, Rachael Robinson, Manasi Raol, Brady Kiesling"
date: "2/14/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::kable
```

# Lab 5

## Goals:

### Identify overall trends in the milk production dataset

### Hypothesize possible explanations for these trends and identify additional data sources that cound be helpful

## Team Section:

### Yearly time series for the milk production (in millions of pounds) by selected States

```{r, echo = FALSE}
milkavg <- milk %>%
  select(state,year,milk_million) %>%
  group_by(year) %>%
  mutate(mavg = mean(milk_million))
milkstate <- milk %>%
  filter(state %in% c('Colorado', 'California', 'Oregon', 'Arizona', 'Pennsylvania')) %>%
  select(state, year, milk_million)
ggplot() +
  geom_point(data = milkstate, aes(x=year, y=milk_million, color = state)) +
  ggtitle("Changes in milk production by state") +
  xlab("Year") +
  ylab("Milk production (millions of pounds)") +
  theme_classic() +
  geom_smooth(data = milkavg, mapping = aes(x=year, y=mavg))
```

This plot indicates that of the 5 states chosen, California has the highest milk production across all years. It can also be noted that while relative increases were seen in the milk production of all the states across the years, California's milk production looks to have increased exponentially over the years. Pennsylvania shows the next largest milk production, with all the other states (Colorado, Oregon, and Arizona) falling to the bottom of the graph. The solid blue line indicates the average milk production for all states over the years. While Arizona, Colorado, and Oregon follow national milk production averages, Pennsylvania and California have historic milk production levels significantly above average. These results seem consistent with a basic assumption that larger states will produce more milk.

### The year when the most milk was produced in the United States
```{r}
library(tidyverse)
milk <- read.csv('state_milk_production.csv')
milk <- milk %>%
  mutate(milk_million = milk_produced/1000000)

m_year=milk %>% group_by(year) %>% summarize(Production=sum(milk_million)) %>% arrange(desc(Production))
m_year[1,]
```
In 2017, 215466 million pounds were produced.

### The year when the least milk was produced in the United States
```{r}
library(tidyverse)
milk <- read.csv('state_milk_production.csv')
milk <- milk %>%
  mutate(milk_million = milk_produced/1000000)

l_year=milk %>% group_by(year) %>% summarize(Production=sum(milk_million)) %>% arrange(Production)
l_year[1,]
```
In 1975, 115398 pounds were produced.

### In 2017, report the 5 states that produced the most milk
```{r}
library(tidyverse)
milk <- read.csv('state_milk_production.csv')
milk <- milk %>%
  mutate(milk_million = milk_produced/1000000)

m_states=milk %>% filter(year==2017) %>% arrange(desc(milk_million))
m_states[1:5,c(2,3,5)]
```
In 2017, California produced 39,798 million pounds, Wisconsin produced 30,320 million pounds, New York produced 14,912 million pounds, Idaho produced 14,627 million pounds, and Texas produced 12,054 million pounds.

### In 2017, report the 5 states that produced the least milk
```{r}
library(tidyverse)
milk <- read.csv('state_milk_production.csv')
milk <- milk %>%
  mutate(milk_million = milk_produced/1000000)

l_states=milk %>% filter(year==2017) %>% arrange(milk_million)
l_states[1:5,c(2,3,5)]
```
In 2017, Alaska produced 3 million pounds, Rhode Island produced 13 million pounds, Hawaii produced 39 million pounds, Arkansas produced 79 million pounds, and Alabama produced 89 million pounds.

### What we found to be the most helpful in learning about this data:

We found it very useful to learn how to manipulate data with filter, select, arrange, and mutate functions. We learned how to effectively create new variables with this data set to both tailor in on specific data (milk production over time for 1 state) and look at the larger data such as the mean milk production across an entire year. The data was particularly interesting because it called attention to state wide differences which would prompt the changes and figures we observed in our plots and calculation. We learned that, on average, larger states produce more pounds of milk annually.


## Individual Section:

### William
```{r}
library(tidyverse)
milk <- read.csv('state_milk_production.csv')
milk <- milk %>%
  mutate(milk_million = milk_produced/1000000)

#Display the distribution of the yearly milk production (in millions of pounds) for 1997
milk1997 <- milk %>%
  filter(year == 1997)

ggplot(data = milk1997, aes(x = milk_million)) +
  geom_density() + 
  ggtitle('Distribution estimate of milk produced in 1997')

#Report the average milk produced by each state
milk1997 %>%
summarize(Average = mean(milk_million))

#Report the median milk produced be each state
milk1997 %>%
summarize(Median = median(milk_million))

#Report the state that produced the most milk
most = milk1997 %>% arrange(desc(milk_million))
most[1,]

#Report the state that produced the least amount of milk
least = milk1997 %>% arrange(milk_million)
least[1,]
```
Above is a density plot of the distribution of milk produced in 1997. The average milk produced this year was 3,121.82 million pounds. The mediam milk produced this year was 1,436 million pounds. The state that produced the most milk this year was California, and they produced 27,582 million pounds. The state that produced the least amount of milk this year was Alaska, and they produced 15 million pounds.

### Rachael

```{r, include=FALSE}

# Load tidyverse
library(tidyverse)
library(knitr)
# Load and view the data
milk <- read.csv('state_milk_production.csv')


milk <- milk %>%
  mutate(milk_million = milk_produced/1000000)

milk1998 <- milk %>%
  filter(year == 1998)


```

```{r, echo = FALSE, warning = FALSE}
ggplot(data = milk1998, mapping = aes(x=milk_million)) +
  geom_density(bins=11, size = 1.5, color = "blue") +
  ggtitle("Distribution of US milk production in 1970") +
  theme_minimal() +
  xlab("Milk in millions of pounds")
```

```{r, include = FALSE}
  meanmilk <- mean(milk1998$milk_million)
  medmilk <- median(milk1998$milk_million)
  meanmilk
  medmilk
```

The plot shows the distribution of milk production in the US for the year 1998. From the plot, it's clear that the density is higher in the lower milk productions, from 0 to 4,000 million pounds. However, it appears there are a few states acting as outliers in the higher milk productions.

When evaluating the data for 1998, I have calcuated the average milk produced across all states to be 3145.22 millions of pounds.

The median was calculated to be 1411.5 millions of pounds.

The large discrepency between the median and the mean, along with the outliers observed, indicates that the mean is skewed upward.

The following is a table representing the top 5 states for milk production in the year 1998:

```{r, results = 'asis', echo = FALSE}

mostmilk <- milk1998 %>%
  select(state, milk_million) %>%
  arrange(desc(milk_million))

kable(mostmilk[1:5,])
```

This indicates that the highest producing state in 1998 was California with 27620 million pounds of milk.

This table shows the 5 states with the lowest milk production in the year 1998:

```{r, results = 'asis', echo = FALSE}

mostmilk <- milk1998 %>%
  select(state, milk_million) %>%
  arrange(desc(milk_million))

kable(mostmilk[45:50,])

```

This indicates that the lowest producing state in 1998 was Alaska with 14 million pounds of milk.


### Brady

```{r echo = TRUE, message = F, warning = F}
milk <- read.csv('state_milk_production.csv')
head(milk)

milk <- milk %>%
  mutate(milk_million = milk_produced/1000000) 

milk1999 <- milk %>%
  filter(year==1999)
milk1999

milkdata<-milk1999$milk_million

ggplot(data=milk1999)+
  geom_density(aes(x=milk_million))



mean(milkdata)
median(milkdata)

most<-milk1999%>%
  arrange(desc(milk_million))
most[1,]

least<-milk1999%>%
  arrange(milk_million)
least[1,]
```
Above I have summarized the data about milk production in the year 1999, the density plot shows that the milk production had a big right skew, meaning a lot of the milk was produced in a few states. This is backed up by the mean value being much higher than the median value. Not suprizingly, alaska produced the least milk, because the climate does not favor dairy production, and California had the most milk production, probably due to the big population and the hospitable climate.

### Manasi
```{r}
library(tidyverse)

milk <- read.csv('state_milk_production.csv')
head(milk)

milk <- milk %>%
  mutate(milk_million = milk_produced/1000000)

milk2000 <- milk %>%
  filter(year == 2000)
milk2000

ggplot(data = milk2000, aes(x = milk_million)) +
  geom_density() + 
  ggtitle('Distribution estimate of milk produced in 2000 by state')

milkdata<-milk2000$milk_million

mean(milkdata)
median(milkdata)

most<-milk2000%>%
  arrange(desc(milk_million))
most[1,]

least<-milk2000%>%
  arrange(milk_million)
least[1,]


```

The plot above shows a density plot of the distribution of milk produced in 2000. The average milk produced was 3347.84 million pounds whereas the median amount of milk produced was 1453.5 million pounds. California produced the most milk in 2000 with about 32,000 million pounds whereas Alaska produced the least with about 13 million pounds. According to my teammates' data this seems to be a trend across many years which makes sense because California has such a large population and Alaska doesn't have a favorable climate when it comes to producing milk.


