---
title: "Bumblebee Lab 6"
date: "2/22/19"
output:
  html_document:
    df_print: paged
authors: Rachael Robinson, William Brickowski, Manasi Raol, Brady Kiesling
---
```{r}
knitr::opts_chunk$set(echo = FALSE)
```



```{r}
library(tidyverse)
COflights <- read_csv("https://files.osf.io/v1/resources/3z5as4pej8/providers/osfstorage/5a8ca28f57103100104584db")
```


# Bumblebee Lab 6

## Research Question: What factors lead to on-time arrivals in Denver?

### Why it's important

  * Why this reason is important?
  
## Supporting Plots

## Conclusions

## Recommendation to DIA Managers

## Individual Plots

### William
```{r}
library(tidyverse)
COflights <- read_csv("https://files.osf.io/v1/resources/3z5as4pej8/providers/osfstorage/5a8ca28f57103100104584db")

OrgDen = COflights %>% filter(LATE_AIRCRAFT_DELAY<=0,CANCELLED!=1,ORIGIN=="DEN")
DestDen = COflights %>% filter(LATE_AIRCRAFT_DELAY<=0,CANCELLED!=1,DEST=="DEN")

OrgDen=OrgDen[c("DAY_OF_MONTH","CARRIER","ORIGIN","DEST","DEP_DELAY","TAXI_OUT",
  "AIR_TIME","DISTANCE","CARRIER_DELAY","WEATHER_DELAY","NAS_DELAY","SECURITY_DELAY")]

DestDen=DestDen[c("DAY_OF_MONTH","CARRIER","ORIGIN","DEST","ARR_DELAY", "TAXI_IN",
  "AIR_TIME","DISTANCE","CARRIER_DELAY","WEATHER_DELAY","NAS_DELAY","SECURITY_DELAY")]

#Given an arrival delay for flights landing in Denver what
#is the probability the flight came from particular airlines?

DestDen_Delayed = DestDen %>% filter(ARR_DELAY>=15)

DestDen_Delayed_Carrier = DestDen_Delayed %>% group_by(CARRIER) %>% count() %>% 
    mutate(Probability = n/dim(DestDen_Delayed)[1])

#Given a departure delay for flights leaving  Denver what
#is the probability the flight came from particular airlines?

OrgDen_Delayed = OrgDen %>% filter(DEP_DELAY >=15)

OrgDen_Delayed_Carrier = OrgDen_Delayed %>% group_by(CARRIER) %>% count() %>% 
    mutate(Probability = n/dim(OrgDen_Delayed)[1])

ggplot(data=DestDen_Delayed_Carrier) +
  geom_bar(aes(x=CARRIER,y=Probability),stat='identity')+
  xlab('Carrier')+
  ylab('Probability')+
  ggtitle('Given an Arrival Delay What is the Probability it Came From a Particular Carrier?')

ggplot(data=OrgDen_Delayed_Carrier) +
  geom_bar(aes(x=CARRIER,y=Probability),stat='identity')+
  xlab('Carrier')+
  ylab('Probability')+
  ggtitle('Given an Departure Delay What is the Probability it Came From a Particular Carrier?')
```
In this code, I divide the data into flight that depart from Denver and those that arrive in Denver in order to isolate the factors that lead to delays in the respective data sets. In the data frame titled OrgDen (flights that originate from Denver), I seek to look at the factors that contribute departure delays. In the data frame titles DestDen (flights whose destination is Denver), I seek to look at the factors that contribute to arrival delays. In my analysis, I look at the conditional probability that given a delay occured, the flight was from a particluar airline. In the cases of both departure delay and arrival, one can see that UA airline has the highest probability of being the airline given that a delay occurar. From this analysis, one can use these result to better understand which airlines contribute to delays in general within the Denver aiport.

### Brady

```{r}

DENflights<-COflights %>%
  filter(DEST=="DEN" ) %>%
  filter(CRS_ARR_TIME>0) %>%
  mutate(isdelayed=ifelse(ARR_DELAY>=15,1,0))


DENflights$CRS_ARR_TIME<-as.numeric(as.character(DENflights$ARR_TIME))

DENflights
  

DENflightsn<-DENflights%>%
  filter(CRS_ARR_TIME<700 | CRS_ARR_TIME>2200)

DENflightsnd<-DENflightsn%>%
  filter(ARR_DELAY>=15)


DENflightsmorning<-DENflights%>%
  filter(CRS_ARR_TIME<=1200 & CRS_ARR_TIME>700)

DENflightsmorning


DENflightsmdelayed<-DENflightsmorning%>%
  filter(ARR_DELAY>=15)




DENflightsa<-DENflights%>%
  filter(CRS_ARR_TIME<1600 & CRS_ARR_TIME>1200)

DENflightsad<-DENflightsa%>%
  filter(ARR_DELAY>=15)


DENflightse<-DENflights%>%
  filter(CRS_ARR_TIME<2200 & CRS_ARR_TIME>1600)

DENflightsed<-DENflightse%>%
  filter(ARR_DELAY>=15)

print("Prortion of night flights delayed")
count(DENflightsnd)/count(DENflightsn)
print("Prortion of morning flights delayed")
count(DENflightsmdelayed)/count(DENflightsmorning)
print("Prortion of afternoon flights delayed")
count(DENflightsad)/count(DENflightsa)
print("Prortion of evening flights delayed")
count(DENflightsed)/count(DENflightse)

ggplot(data=DENflights)+
  geom_histogram(aes(x=CRS_ARR_TIME, fill=as.factor(isdelayed)), bins=24)+
  scale_fill_discrete(name="Is Delayed", labels=c("Not Delayed","Delayed","NA"))
```
In this code, I looked into how the scheduled landing time in Denver affected whether the plane was delayed or not. I groped the day into 4 sections, Night, Morning, Evening and Afternoon and calculated the probability the flight was delayed given the flight was scheuled to land at that time(P(delayed|time of day)).I found that flights at night were the most likely to be delayed(10pm-7am) while flights(7am-noon) were the least likely to be delayed. I made a stacked histogram showing the time of day the flight was schedueld, and had a one color the flights that arrived on time and one color the flights that were delayed. You can clearly see the increase in the porportion of delayed flights from mornign until evening, but becuase there are so few data samples from the night, It is hard to tell the proportion of delayed flights. My theory why there are more delayed flights in the eveing, is that things get slowly backed up over the course of the day. 

### Manasi

```{r}
library(tidyverse)
COflights <- read_csv("https://files.osf.io/v1/resources/3z5as4pej8/providers/osfstorage/5a8ca28f57103100104584db")

filter(COflights, DEST=="DEN" & ARR_DELAY >=15 & !is.na(ARR_DELAY) & WEATHER_DELAY >= 15 & CARRIER_DELAY >= 15 & NAS_DELAY >= 15 & SECURITY_DELAY >= 15 & !is.na(WEATHER_DELAY) & !is.na(CARRIER_DELAY) & !is.na(NAS_DELAY) & !is.na(SECURITY_DELAY))

ggplot(data = COflights, aes(x = WEATHER_DELAY)) +
  geom_dotplot() +
  ggtitle('Delay in Minutes Due to Weather') +
  labs(x="Minutes of Delay", y="Number of Delayed Flights") 

ggplot(data = COflights, aes(x = CARRIER_DELAY)) +
  geom_dotplot() +
  ggtitle('Delay in Minutes Due to Carrier') +
  labs(x="Minutes of Delay", y="Number of Delayed Flights")

ggplot(data = COflights, aes(x = NAS_DELAY)) +
  geom_dotplot() +
  ggtitle('Delay in Minutes Due to NAS') +
  labs(x="Minutes of Delay", y="Number of Delayed Flights")

ggplot(data = COflights, aes(x = SECURITY_DELAY)) +
  geom_dotplot() +
  ggtitle('Delay in Minutes Due to Security') +
  labs(x="Minutes of Delay", y="Number of Delayed Flights")

```

I created these plots in order to determine out of all the provided reasons for delays, which one was the most common and caused the longest delays. Based on the plots, the longest delays as well as the most delays were due to specific carriers. However, the second most common reason for delay was due to issues with the National Air System. This is something that can be solved with better scheduling and solutions that produce less technical issues. This shows that delayed flights into Denver have the most issues with the carrier and problems with the National Air System. 

### Rachael

```{r}

library(tidyverse)
COflights <- read_csv("https://files.osf.io/v1/resources/3z5as4pej8/providers/osfstorage/5a8ca28f57103100104584db")

flightnotesA <- COflights %>%
  filter(DEST == "DEN", !is.na(ARR_TIME), !is.na(AIR_TIME), !is.na(ARR_DELAY)) %>%
  select(ARR_DELAY,AIR_TIME,DISTANCE) %>%
  mutate(DELAY = ifelse(ARR_DELAY >= 15, 1, 0)) %>%
  group_by(AIR_TIME) %>%
  filter(sum(DELAY == 0, DELAY == 1) > 30) %>%
  summarise(timeprop = mean(DELAY))



flightnotesD <- COflights %>%
  filter(ORIGIN == "DEN", !is.na(ARR_TIME), !is.na(AIR_TIME), !is.na(ARR_DELAY)) %>%
  select(ARR_DELAY,AIR_TIME,DISTANCE) %>%
  mutate(DELAY = ifelse(ARR_DELAY >= 15, 1, 0)) %>%
  group_by(AIR_TIME) %>%
  filter(sum(DELAY == 0, DELAY == 1) > 30) %>%
  summarise(timeprop = mean(DELAY))


ggplot() +
  geom_line(data = flightnotesA, mapping = aes(x=AIR_TIME, y=timeprop, color = "Destination")) +
  geom_line(data = flightnotesD, mapping = aes(x=AIR_TIME, y=timeprop, color = "Origin")) +
  labs(title = "Probability of a delayed landing at various air times", subtitle = "DEN airport as the origin/destination", x = "Air time (minutes)", y = "Probability of delayed (>15 min) landing", color = "DEN") +
  theme_classic()

```
Finding #1- To manipulate this data I first looked at flights arriving to Denver (i.e. Destination DEN) and calculated the proportion of flights with delayed arrival, corresponding to their relative flight time. In order to rule out probabilistic outliers, I referenced the CLT and only calculated proportions of delayed arrival for air times with more than 30 data points. However, there are clearly still some high leverage points remaining. For flights not unusually long, the plot shows a positive trend between air time and probability of delayed arrival. From around 20 to 280 minutes, an increase in probability in demonstrated. This implies that a longer flight (while not unusually long) is more likely to arrive to DEN delayed.

Finding #2- I manipulated the data in a similar way to calculate the proportion of delayed arrivals where DEN is the origin airport. Outliers were regulated for in the same way as before. However, the positive trend is still the same. As air time increases, the probability of having a delayed arrival at a different airport increases. This implies that a longer flight (while not unusually long) is more likely to arrive at other airports from DEN, delayed.

  
```{r, include = FALSE}
flightnormal <- COflights %>%
  filter(AIR_TIME == 104) %>%
  select(ARR_DELAY,AIR_TIME,DISTANCE)

shapiro.test(flightnormal$ARR_DELAY)
  
  
```

## Individual contribution and non-defaults in plot generation

### William

### Brady

### Manasi

### Rachael

