---
title: "Lab8_Bumblebee"
author: "Rachael Robinson, William Brickowski, Manasi Raol, Brady Kiesling"
date: "3/5/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Lab 8

```{r}
#general tidying of data, selecting relevant variables, differentiating date and time and converting them from characters

library(tidyverse)
ufo <- read_csv("ufo_sighting_data.csv")
ufo <- ufo %>% select(-described_duration_of_encounter, -description) %>%
  separate(Date_time, into = c("Date","Time"), sep = " ")
ufo$Date <- as.Date(ufo$Date, "%m/%d/%Y")
ufo$Time <- parse_time(ufo$Time, format = "%H:%M")

#ufo is tidy

```


## Overall Team Question: What factors correlate with UFO sightings, and what variables could account for variation in sighting duration?

## Why this question is important

  * why this reason is important
  
## Conclusion/Answer to question

## Recommendation/importance to domain expert

## Description of datasets imported/tidied/analyzed

## Individual findings

### William

### Brady

### Manasi

### Rachael

## Who did what/tidyverse functions used

### William

### Brady

### Manasi

### Rachael
