---
title: "Lab8_Bumblebee"
author: "Rachael Robinson, William Brickowski, Manasi Raol, Brady Kiesling"
date: "3/5/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Lab 8

```{r}
#general tidying of data, selecting relevant variables, differentiating date and time and converting them from characters

library(tidyverse)
ufo <- read_csv("ufo_sighting_data.csv")
ufo <- ufo %>% select(-described_duration_of_encounter, -description) %>%
  separate(Date_time, into = c("Date","Time"), sep = " ")
 ufo$Date <- as.Date(ufo$Date, "%m/%d/%Y")
 ufo$Time <- parse_time(ufo$Time, format = "%H:%M")

#ufo is tidy

```


## Overall Team Question: What factors correlate with UFO sightings, and what variables could account for variation in sighting duration?

Our domain expert is Judy Messoline, founder and owner of the UFO WatchTower located in Hooper, Colorado. As an avid consumer of all things alien, Messoline would naturally be curious about the circumstances which prompt more UFO and alien sightings. Messoline might also be interested in creating an exicting atmosphere for watch tower visitors, so she may find our results interesting to incorporate new elements to her business, and drive her profit. Finally, Messoline might be interested in having longer UFO sightings herself, so finding out what variables increase duration would be interesting to her.

## Why this question is important

This question is important first because of the little information known about UFO sightings. Analysis of data is crucial to understand what influences UFO sightings, reportings, and all the time in between. This question is also important because of prejudices associated with those who report UFO sightings. With concrete data analysis, this question helps illuminate some of the unifying circumstances which prompt individuals to report UFO sightings, and possibly unifying characteristics of those people.

  * This reason is important because we know so little about the possibilities of extraterrestrial life. Analysis of this data is how we can demystify some of the circumstances surrounding alien sightings, and could lead to discovery of patterns. Finding those patterns would be the next step to recreating circumstances and developing tests.
  
## Conclusion/Answer to question


## Recommendation/importance to domain expert


## Description of datasets imported/tidied/analyzed

The dataset we have chosen to import, tidy, and analyze can be found at: https://www.kaggle.com/camnugent/ufo-sightings-around-the-world.

This data set contains over 80,000 reports of UFO sightings. The data includes the date and time the UFO was seen, the location the UFO was observed, a description of the UFO, the sighting, and its duration, and the date the sighting was reported. We were able to tidy the dataset in order to accurately represent when and where the UFO was seen. We tidied the data to primarily focus on the location, time, and duration of the UFO sighting. Using this information we were able to answer our central question as to what factors influenced a UFO sighting and it's duration.

## Individual findings

### William


```{r, echo = FALSE, message = FALSE, error= FALSE}
library(tidyverse)
ufo = read_csv("ufo_sighting_data.csv")

ufo = ufo %>% select(1,6,9:11) %>% drop_na() %>% separate(1, c('Date','Time'), sep = " ", remove = TRUE) 

ufo=as_tibble(ufo)
ufo$Time=parse_number(ufo$Time)
ufo$Time=parse_integer(ufo$Time)
ufo$latitude=parse_double(ufo$latitude)
ufo$longitude=parse_double(ufo$longitude)
ufo$date_documented=parse_date(ufo$date_documented,format="%m/%d/%Y")
ufo$Date=parse_date(ufo$Date,format="%m/%d/%Y")

ufo = ufo %>% mutate(diff=date_documented-Date,day=format(ufo$Date, "%d"),month=format(ufo$Date, "%m"),year=format(ufo$Date, "%Y"))

ufo$diff=parse_number(ufo$diff)
ufo$day=parse_integer(ufo$day)
ufo$month=parse_integer(ufo$month)
ufo$year=parse_integer(ufo$year)

lmod=lm(length_of_encounter_seconds ~ Time + day + month + year + diff + latitude + longitude, data=ufo)
summary(lmod)
```
In my analysis of the ufo sighting data set, I sought to use multivariate linear regression with the length of the encounter of a ufo sighting as the response variable and with the following predictors: time of day, day of the week, month, year, number of days between when the ufo was sighted (represented by the column titled 'diff') and when its sighting was recorded, latitude of where it was observed, and the respective longitude. After performing multivariate regression on the data, I found that at the 1% confidence level the only significant predictors was 'diff' as it had an associated P-value of 0.00796. The beta associated with this predictor was 4.139 which means that on average if you increase the number of days between when the sighting occured and when it was recorded by 1, the sighting increases by 4.139 seconds. This result makes sense because people, psychologically speaking, tend to overestimate emotional events due to the flashbulb memory theory. However, with time, it has been shown that humans exaggerate what they had seen. The data is showing that this is most likely variable to determine the duration of a sighting.


### Brady

### Manasi
```{r}
ufo_times <- ufo %>%
  separate(Time, into = c("Hour","Minutes","Seconds"), sep = ":")

ufo_comb_times <- ufo_times %>%
  unite(Time, Hour, Minutes, Seconds, sep = "")

ufo_time_data <- ufo_comb_times %>%
  group_by(Time)

```

```{r}
ggplot(data = ufo_comb_times, aes(x = Time, y = count)) +
  geom_jitter()
```
Findings:


### Rachael

```{r}
library(hms)
library(lubridate)

ufo_days <- ufo %>%
  mutate(dayofweek = weekdays(ufo$Date))

ufo_sum <- ufo_days %>%
  group_by(dayofweek) %>%
  summarise(daycount = n())


```

```{r}
ggplot(data = ufo_sum, aes(x=dayofweek, y= daycount, fill = dayofweek)) +
  geom_bar(stat = "identity") +
  scale_x_discrete(limits = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")) +
  coord_cartesian(ylim = c(5000,14000)) +
  labs(title = "Number of UFO Sightings by Day of Week", x = "Day of the week", y = "Number of sightings", fill = "Day of week") +
  theme_classic()
  

```

Findings: The above plot suggests that day of the week does correlate with number of reported UFO sightings. From the beginning of the week (Monday), UFO sighting reports steadily increase throughout the week, creating a significant peak on Saturday. The data suggests that a UFO is most likely to be seen on a Saturday, and least likely to be seen on a Monday, and a steady increase in between. Sunday reportings are comparable with Friday, and do represent the decline from high reporting Saturday to the dropoff on Monday.

There could be many reasons for this correlation, including some confounding variables not represented in the data set. For example, individuals might be more likely to be outside to see a UFO on the weekend. Alternatively, individuals might also be more likely to consume substances on the weekend, which might increase the number of preported sightings.

Ultimately, it is clear that there is some relationship between day of the week and UFO sightings, particularly that the weekend is a popular time for UFO sightings, where Monday is relatively unpopular for UFO sightings.


## Who did what/tidyverse functions used

### William


In order to perform multivariate regression, there were several things that had to occur prior. I had to clean and tidy the data in an way that would permit regression, as only numerical values can be regressed against each other. To begin with, I loaded the data set by using the function read_csv. I then, using tidyverse and the pip function, selected only those columns that would be usefull for my analysis. In this case, the columns were Date, Time, length_of_encounter_seconds, date_documented, latitude, and longitude. I then split the Date column into two columns because it originally contained both the date and time. I called the sepearte function on the Date column with " " as the seperater. I then convereted all columns to numeric or integer values. By converting the Time column to a numberic it gave me just the hour which is what I wanted to use to regress. I also parsed the Date and date_documented columns to date using the parse_date function. Then using the mutate function, I created several columns called diff, day, month, and year - all of which would be used as predictors in the model. The diff column is difference between the date_documented and Date. This returns are char so I had to convert this column to a numeric value later on. As for the day, month, and year columns, I created these by using the format function and selecting the respective value from the original Date column. These all had to be converted to numeric afterwards as well. With all of my data in the right format, I then ran linear regresson with length_of_encounter_seconds as the response and Time, day, month, year, diff, latitude, and longitude as predictors. Finally, the summary function was called in order to show the significance of the respective predictors as well as their associated betas.


### Brady

### Manasi
In this lab, I helped tidy the data using functions such as separate to make sure the dataset followed tidying rules. For example, the original data had the date and time in one column, but I used separate to create separate columns for the date and time variables. 
For my individual part, I manipulated the data to see if there were more UFO sitings at specific times. First, I separated the Time variable into hour, minutes, and seconds by the colons and then united them to form a six digit time for each siting to make the values easier to plot. This way I would know that a higher number such as 221500 (22:15:00) is time that is later in the day whereas a lower number such as 061000 (06:10:00) is earlier in the day.I then used ggplot to plot the time vs the total number of sitings to see which times of day had the most sitings. I added lables and a title to make my graph clearer.

### Rachael

For this lab, I searched for data sets and found the data set we eventually used for the lab. I wrote the code to import the data, selected the relevant variables, and converted the data and time to appropriate classes instead of the default string. I also wrote about our domain expert, the importance/relevance of the question, and the explanation of the dataset.

For my individual work, I manipulated the data further to create a new variable representing the day of the week the UFO was seen. I then summarised the data to find the total counts of UFO sightings for each day of the week. Finally, I used ggplot to visualize this data. Within this visualization I used the fill command to color by day, the scale function to reorder the bars to move in a chronological order, and a coordinate switch to change the limits of the y axis. Finally, I used the lab and theme functions to improve the aesthetics and clarity of my plot.
