---
title: "Bumblebee Lab 9"
author: "Rachael Robinson, William Brickowski, Manasi Raol, Brady Kiesling"
date: "3/13/2019"
output: html_document
---

#Lab 9

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(lubridate)
library(OpenStreetMap)
```

```{r}
#global data set, make duplicates before manipulation when working on individual portions

station <- read_csv("Nice_Ride_2017_Station_Locations.csv")
trip <- read_csv("Nice_ride_trip_history_2017_season.csv")

```

##Team Question:
Are more rides taken where the concentration of nice ride stations is higher?

##Most interesting plot to answer team question:

```{r}
autoplot(map.latlon) +  
  geom_point(data=locations, 
               aes(x=Longitude, y=Latitude), 
               color = 'red', size = 0.8) +
  labs(x='Longitude', y='Latitude') +
  ggtitle('Stations Across Minneapolis and Surrounding Areas')
```



##Individual Findings

###William

```{r}
locations=read_csv("Nice_Ride_2017_Station_Locations.csv")
history=read_csv("Nice_ride_trip_history_2017_season.csv")

history=rename(history,Number = `Start station number`)

history = history %>% 
  left_join(locations, by = "Number")

history=rename(history, `Start station number` = `Number`, `Number` = `End station number`,`Start station name`=`Name`,`Start station latitude`=`Latitude`,`Start station longitude`=`Longitude`,`Start station total docks`=`Total docks`)

history = history %>% 
  left_join(locations, by = "Number")

history=rename(history, `End station number` = `Number`,`End station name`=`Name`,`End station latitude`=`Latitude`,`End station longitude`=`Longitude`,`End station total docks`=`Total docks`)

history=history %>% select(1:8,10:12,14:16)

history = history %>% mutate(trip_name=str_c(`Start station`, `End station`, sep = " to "))

history = history %>% drop_na() %>% separate(1, c('Start date','Start time'), sep = " ", remove = TRUE) 

history = history %>% separate(5, c('End date','End time'), sep = " ", remove = TRUE)

most_freq = history %>% count(trip_name) %>% arrange(desc(n))

head(most_freq)
```

Above is a table of the top 6 most frequent rides in the data set. The five most frequent rides have the same start and end station. For example, the most frequent ride with a total of 4853 rides is from Lake Street & Knox Ave S and back to Lake Street & Knox Ave S. This is followed a total of 3171 rides from Lake Calhoun Center and back to Lake Calhoun Center. This data is very interesting because it can be used to understand how consumers use this ride serivce. From the data, it is aparent that most individuals start and return to the same station. This is vital data because while particular stations may be more popular than others, it must be taken into account that vehicles will be return to these stations. So when determining how many docks are needed at popular locations, the provider must not only consider the most popular stations but also how frequent customers return scooters to these locations. If a particular location is a common one in which vehicles return, then less docks will be needed. The rate by which scooters return to a particular location can be modeled by using a poisson process in order to maximize efficiency. The sixth most popular ride is one that deviates from this pattern as the start and end location are different. There were 1857 rides from Willey Hall to Weisman Art Museum. In this case, this trip may represent a destination trip rather than a return trip. By classyfing particular rides as desintaiton (one in which the scooter starts somewhere and is dropped off somewhere else) or return trips (one in which a customer starts and ends at the same location), the provider can again better estimate how many docks are needed at eaach location.

###Brady


###Manasi

```{r}
library(tidyverse)
library(lubridate)
library(OpenStreetMap)

station <- read_csv("Nice_Ride_2017_Station_Locations.csv")
trip <- read_csv("Nice_ride_trip_history_2017_season.csv")

christmas <- station %>%
  filter(Startingdate == "2017-12-25") 

autoplot(map.latlon) + 
  geom_segment(data=christmas, 
               aes(x=start.long, y=start.lat,
                   xend=end.long, yend=end.lat), 
               color = 'green', size = .6,
               arrow = arrow(length = unit(0.2, "cm"))) +
  labs(x='Longitude', y='Latitude') +
  scale_x_continuous(limits = c(-93.3, -93.2)) +
  scale_y_continuous(limits = c(44.93, 45.02)) +
  ggtitle('Christmas Rides')
```

```{r}
library(tidyverse)
library(lubridate)
library(OpenStreetMap)

station <- read_csv("Nice_Ride_2017_Station_Locations.csv")
trip <- read_csv("Nice_ride_trip_history_2017_season.csv")

thanksgiving <- station %>%
  filter(Startingdate == "2017-11-23") 


autoplot(map.latlon) + 
  geom_segment(data=thanksgiving, 
               aes(x=start.long, y=start.lat,
                   xend=end.long, yend=end.lat), 
               color = 'orange', size = .6,
               arrow = arrow(length = unit(0.2, "cm"))) +
  labs(x='Longitude', y='Latitude') +
  scale_x_continuous(limits = c(-93.3, -93.2)) +
  scale_y_continuous(limits = c(44.93, 45.02)) +
  ggtitle('Thanksgiving Rides')
```

I looked at rides on popular holidays (Thanksgiving and Christmas) to see if there were more rides or less rides on those days. This was in order to determine whether more bikes were needed during the holidays or not. According to the Minneapolis bike maps for Thanksgiving and Christmas, there are less riders during these two holidays. This makes sense because most people stay in with their families during these times of the year. Additionally, these two holidays are during the winter season so less people may want to bike places due to the cold weather (especially in Minneapolis). 

###Rachael

```{r}
#setup

library(tidyverse)
library(lubridate)
library(OpenStreetMap)

#loading the data

station_loc <- read_csv("Nice_Ride_2017_Station_Locations.csv")
Nice_Trip <- read_csv("Nice_ride_trip_history_2017_season.csv")

#simplifying the station data

station_loc <- station_loc %>%
  select(-Name, -`Total docks`)

#joining data sets, renaming variables

Nice_Trip <- Nice_Trip %>%
  left_join(station_loc, c("Start station number" = "Number"))

Nice_Trip <- Nice_Trip %>%
  rename(start_lat = Latitude, start_long = Longitude)

Nice_Trip <- Nice_Trip %>%
  left_join(station_loc, c("End station number" = "Number"))

Nice_Trip <- Nice_Trip %>%
  rename(end_lat = Latitude, end_long = Longitude)


#creating subsets, longest/shortest rides

longest_trip <- Nice_Trip %>%
  top_n(1000, `Total duration (Seconds)`)

shortest_trip <- Nice_Trip %>%
  top_n(-1000, `Total duration (Seconds)`)

#Member data frames on longest/shortest rides

member_count_long <- longest_trip %>%
  count(Account = `Account type`)

member_count_short <- shortest_trip %>%
  count(Account = `Account type`)

#Data frame showing discrepency, casual riders more represented in starting and ending at the same station

casual_trip <- Nice_Trip %>%
  filter(`Start station` == `End station`) %>%
  count(Account = `Account type`)

#Renaming values in data frame

member_count_long <- member_count_long %>%
  rename(`Long ride count` = n)

member_count_short <- member_count_short %>%
  rename(`Short ride count` = n)

causal_trip <- casual_trip %>%
  rename(`Ride count` = n)

#Formatting date in datasets, 

longest_trip <- longest_trip %>%
  separate(`Start date`, into = c("Start_Date","Time"), sep = " ")

shortest_trip <- shortest_trip %>%
  separate(`Start date`, into = c("Start_Date","Time"), sep = " ")

longest_trip$Start_Date <- as.Date(longest_trip$Start_Date, "%m/%d/%Y")
shortest_trip$Start_Date <- as.Date(shortest_trip$Start_Date, "%m/%d/%Y")

#Added variable indicating the weekday of the ride on both data frames

daytrip_long <- longest_trip %>%
  mutate(dayofweek = weekdays(Start_Date)) %>%
  count(day_of_week = dayofweek)

daytrip_short <- shortest_trip %>%
  mutate(dayofweek = weekdays(Start_Date)) %>%
  count(day_of_week = dayofweek)

#renaming columns in weekday info, joining tables for short and long rides

daytrip_long <- daytrip_long %>%
  rename(long_trip_count = n)

daytrip_short <- daytrip_short %>%
  rename(short_trip_count = n)

daytripper <- left_join(daytrip_long, daytrip_short, key = "day_of_week")

#data frame for map plot

long_map <- longest_trip %>%
  top_n(10, `Total duration (Seconds)`)

short_map <- shortest_trip %>%
  top_n(-10, `Total duration (Seconds)`)

```

The following tables show the breakdown of member vs casual riders in the top 1000 longest and 1000 shortest rides in the data set. Note that the short ride count totals to more than 1000- this is because many of the short ride durations are equal, and the 1000 shortest rides refers to the 1000 shortest ride durations.

```{r}
print(member_count_long)
print(member_count_short)
```

One finding from these tables is that the longest rides in the set were primarily from casual riders, whereas the opposite is true for the shortest rides in the set. Thus, it may be true that shorter rides are more likely to be member riders, and longer rides are more likely to be casual riders.

One discrepency exists in short rides which start and end at the same station. Upon inspection of the data, it appears that casual riders are twice as likely to start and end at the same station. Ruling these cases out in the shortest ride data presents a bias against casual riders in the data representation.

```{r}
print(casual_trip)
```

Next, I investigated the days of the week the longest and shortest rides were happening on. The counts are summarised in the following chart.

```{r}
ggplot(data = daytripper, aes(fill = day_of_week)) +
  geom_bar(mapping = aes(x=day_of_week, y=long_trip_count), stat = "identity") +
  scale_x_discrete(limits = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")) +
  geom_point(mapping = aes(x=day_of_week, y=short_trip_count)) +
  geom_line(mapping = aes(x=day_of_week, y=short_trip_count, group = 1)) +
  labs(title = "Weekday distribution of longest and shortest rides", subtitle = "Bars represent long trips, line shows short trips", x = "Day of week", y = "Number of rides out of shortest/longest", fill = "Day of week")
```

The bars show the distribution of the longest rides in the data set. Clearly, the longest rides are more likely to happen on weekend days, with a prominent peak on Saturday. This is congruent with our intuition that individuals would be more likely to take longer rides on the weekend. Similarly, the shortest rides in the data set (represented by the black lines) are much more consistent across the work week, and decline slightly on the weekends.

We can conclude from this chart that the longest rides in the dataset occur on the weekends, whereas the shortest rides in the dataset are more evenly distributed throughout the week, with a small decline on the weekends.

Finally, we can plot the 10 longest and 10 shortest rides over a map of Minneapolis.

```{r}
# Set latitudes and longitudes of city map
LAT1 <- 44.88 
LAT2 <- 45.05
LON1 <- -93.35
LON2 <- -93.08

# Generate map
map <- openmap(c(LAT2,LON1), c(LAT1,LON2), zoom = NULL,
               type = "esri",
               mergeTiles = TRUE)

# Project map to latitude and longitude
map.latlon <- openproj(map, projection = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")

# Plot map of Minneapolis
autoplot(map.latlon) +
  geom_segment(data = long_map, aes(x=start_long, y=start_lat, xend = end_long, yend = end_lat), color = 'blue', size = .75, arrow = arrow(length = unit(0.2, "cm"))) +
  geom_point(data = short_map, aes(x=start_long, y=start_lat), color = 'red', size = 2.5) +
  geom_point(data = short_map, aes(x=end_long, y=end_lat), color = 'yellow', size = .75, shape = 8) +
  labs(title = "Top 10 longest and shortest bike rides in Minneapolis, 2017", subtitle = "Blue arrows represent longest trips, red indicates short ride start stations, and yellow stars are short ride end stations", x = "Longitude", y = "Latitude")
```

Again, it's easy to see how the short rides often start and end in the same location. Furthermore, most of the shortest rides occur near bodies of water in the area, and are generally clustered toward the center of the city. On the other hand, the longest rides in the set unsurprisingly cover larger distances, and are less centralized. While some of the distances are relatively short, the longer rides on this plot are generally farther apart.

This plot is illuminating to show the popular locations of the shortest and longest rides in the set.


## Who did what/tidyverse functions used


###William
For this lab, I sought to anser the question of which trips are the most popular in order to learn more about how many docks should be at each location. In order to perform this data analysis, I had to make use of several tidyverse functions. I initially used the read_csv function to import the respective data sets. I had to first rename the Start station number from one of the tibbles to Number so that this column title would match so that I could perform a left join. I joined the data sets based on the Number of the ride. I used left join because I only wanted longitudes and latitudes for rides the stations that had rides occur. I then performed another left join in order to get latidude and longitue for the End station. I then selected only those columsn needed for my analysis. I mutated my new tibble to get a new column for trip_name by concatenating Start station and End station. I then dropped the na values. And, I seperated the Start and End date columns so that I got two columns for each corresponding to Date and Time. Finally, I used the pip to count the trip_name column and arrange in descending order of n to show the most frequent trips.

###Brady


###Manasi

For this lab, I answered the question of whether more or less bikes are needing during the holidays. I created maps with arrows repesenting peoples' bike paths during Thanksgiving and Christmas which are two of the most popular holidays. I customized the maps with specific arrows, colors, labels, and axes. I used filter to create the specific maps for Thanksgiving and Christmas and used autoplot to create the maps. I also created the team plot to answer the overall team question. I found that more stations are concentrated within the hearts of the cities such as Minneapolis and St. Paul which makes sense since their populations are higher. Accoring to findings made my group, there are a couple streets within these cities that get the most rides. Therefore, a possible solution may be to add more bikes to these stations so more people can take advantage of them.


###Rachael

For this lab, I created and formatted the team document. I also worked to override defaults on my computer in order to load and run the OpenStreetMap library to compile the document for my group. In addition, I completed my individual analysis and reported several findings about the data. In the tidyverse, I read in the data and completed some mutating joins to combine data from the two tables. I used the select, filter, rename functions to simplify my data frames. I used the mutate, top_n(), and count functions to create unique variables of which I could evaluate the characteristics of the longest and shortest rides in the dataset. Finally, I used functions such as scale_x_discrete() and labs() to customize my charts and plots. 

