---
title: "Untitled"
author: "Rachael Robinson"
date: "3/14/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(lubridate)
library(OpenStreetMap)
```

```{r}
station_loc <- read_csv("Nice_Ride_2017_Station_Locations.csv")
Nice_Trip <- read_csv("Nice_ride_trip_history_2017_season.csv")
```
```{r}
head(station_loc)
head(Nice_Trip)
```

```{r}
station_loc <- station_loc %>%
  select(-Name, -`Total docks`)

head(station_loc)
```

```{r}
Nice_Trip <- Nice_Trip %>%
  left_join(station_loc, c("Start station number" = "Number"))

head(Nice_Trip)
```

```{r}
Nice_Trip <- Nice_Trip %>%
  rename(start_lat = Latitude, start_long = Longitude)

head(Nice_Trip)

```

```{r}
Nice_Trip <- Nice_Trip %>%
  left_join(station_loc, c("End station number" = "Number"))

head(Nice_Trip)
```

```{r}
Nice_Trip <- Nice_Trip %>%
  rename(end_lat = Latitude, end_long = Longitude)

head(Nice_Trip)
```

```{r}
longest_trip <- Nice_Trip %>%
  top_n(1000, `Total duration (Seconds)`)

shortest_trip <- Nice_Trip %>%
  top_n(-1000, `Total duration (Seconds)`)

#view(shortest_trip)

```

Data analysis 

Members/casual

```{r}
member_count_long <- longest_trip %>%
  count(... = `Account type`)

print(member_count_long)
```
```{r}
member_count_short <- shortest_trip %>%
  count(... = `Account type`)

print(member_count_short)
```

A side note about why I'm keeping more data points in the casual riders.
(Tried to rule out instances where the start and end stations were the same, realized it was a biasing results and hiding casual riders taking short trips)


```{r}
casual_trip <- Nice_Trip %>%
  filter(`Start station` == `End station`) %>%
  count(... = `Account type`)

print(casual_trip)

```
An explanation for the discrepency I found when ruling out really short rides with the same start and end station - casual riders are more likely to start and end at the same station at over a 2 to 1 ratio.





Day of week breakdown

```{r}
longest_trip <- longest_trip %>%
  separate(`Start date`, into = c("Start_Date","Time"), sep = " ")

shortest_trip <- shortest_trip %>%
  separate(`Start date`, into = c("Start_Date","Time"), sep = " ")

longest_trip$Start_Date <- as.Date(longest_trip$Start_Date, "%m/%d/%Y")
shortest_trip$Start_Date <- as.Date(shortest_trip$Start_Date, "%m/%d/%Y")

head(longest_trip)
head(shortest_trip)

```



```{r}
daytrip_long <- longest_trip %>%
  mutate(dayofweek = weekdays(Start_Date)) %>%
  count(day_of_week = dayofweek)

daytrip_short <- shortest_trip %>%
  mutate(dayofweek = weekdays(Start_Date)) %>%
  count(day_of_week = dayofweek)

print(daytrip_long)
```

```{r}
print(daytrip_short)
```

```{r}
daytrip_long <- daytrip_long %>%
  rename(long_trip_count = n)

daytrip_short <- daytrip_short %>%
  rename(short_trip_count = n)


daytripper <- left_join(daytrip_long, daytrip_short, key = "day_of_week")

  
print(daytripper)
```

```{r}
ggplot(data = daytripper, aes(fill = day_of_week)) +
  geom_bar(mapping = aes(x=day_of_week, y=long_trip_count), stat = "identity") +
  scale_x_discrete(limits = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")) +
  geom_point(mapping = aes(x=day_of_week, y=short_trip_count)) +
  geom_line(mapping = aes(x=day_of_week, y=short_trip_count, group = 1)) +
  labs(title = "Weekday distribution of longest and shortest rides", subtitle = "Bars represent long trips, line shows short trips", x = "Day of week", y = "Number of rides out of shortest/longest", fill = "Day of week")
```

```{r}
long_map <- longest_trip %>%
  top_n(10, `Total duration (Seconds)`)

short_map <- shortest_trip %>%
  top_n(-10, `Total duration (Seconds)`)
```




```{r}
# Set latitudes and longitudes of city map
LAT1 <- 44.88 
LAT2 <- 45.05
LON1 <- -93.35
LON2 <- -93.08

# Generate map
map <- openmap(c(LAT2,LON1), c(LAT1,LON2), zoom = NULL,
               type = "esri",
               mergeTiles = TRUE)

# Project map to latitude and longitude
map.latlon <- openproj(map, projection = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")

# Plot map of Minneapolis
autoplot(map.latlon) +
  geom_segment(data = long_map, aes(x=start_long, y=start_lat, xend = end_long, yend = end_lat), color = 'blue', size = .75, arrow = arrow(length = unit(0.2, "cm"))) +
  geom_point(data = short_map, aes(x=start_long, y=start_lat), color = 'red', size = 2.5) +
  geom_point(data = short_map, aes(x=end_long, y=end_lat), color = 'yellow', size = .75, shape = 8) +
  labs(title = "Top 10 longest and shortest bike rides in Minneapolis, 2017", subtitle = "Blue arrows represent longest trips, red indicates short ride start stations, and yellow stars are short ride end stations", x = "Longitude", y = "Latitude")
```


```{r}
#setup

library(tidyverse)
library(lubridate)
library(OpenStreetMap)

#loading the data

station_loc <- read_csv("Nice_Ride_2017_Station_Locations.csv")
Nice_Trip <- read_csv("Nice_ride_trip_history_2017_season.csv")

#simplifying the station data

station_loc <- station_loc %>%
  select(-Name, -`Total docks`)

#joining data sets, renaming variables

Nice_Trip <- Nice_Trip %>%
  left_join(station_loc, c("Start station number" = "Number"))

Nice_Trip <- Nice_Trip %>%
  rename(start_lat = Latitude, start_long = Longitude)

Nice_Trip <- Nice_Trip %>%
  rename(end_lat = Latitude, end_long = Longitude)

Nice_Trip <- Nice_Trip %>%
  rename(end_lat = Latitude, end_long = Longitude)


#creating subsets, longest/shortest rides

longest_trip <- Nice_Trip %>%
  top_n(1000, `Total duration (Seconds)`)

shortest_trip <- Nice_Trip %>%
  top_n(-1000, `Total duration (Seconds)`)

#Member data frames on longest/shortest rides

member_count_long <- longest_trip %>%
  count(Account = `Account type`)

member_count_short <- shortest_trip %>%
  count(Account = `Account type`)

#Data frame showing discrepency, casual riders more represented in shorter rides

casual_trip <- Nice_Trip %>%
  filter(`Start station` == `End station`) %>%
  count(Account = `Account type`)

#Renaming values in data frame

member_count_long <- member_count_long %>%
  rename(count = n)

member_count_short <- member_count_short %>%
  rename(count = n)

causal_trip <- casual_trip %>%
  rename(count = n)

#Formatting date in datasets, 

longest_trip <- longest_trip %>%
  separate(`Start date`, into = c("Start_Date","Time"), sep = " ")

shortest_trip <- shortest_trip %>%
  separate(`Start date`, into = c("Start_Date","Time"), sep = " ")

longest_trip$Start_Date <- as.Date(longest_trip$Start_Date, "%m/%d/%Y")
shortest_trip$Start_Date <- as.Date(shortest_trip$Start_Date, "%m/%d/%Y")

#Added variable indicating the weekday of the ride on both data frames

daytrip_long <- longest_trip %>%
  mutate(dayofweek = weekdays(Start_Date)) %>%
  count(day_of_week = dayofweek)

daytrip_short <- shortest_trip %>%
  mutate(dayofweek = weekdays(Start_Date)) %>%
  count(day_of_week = dayofweek)

#renaming columns in weekday info, joining tables for short and long rides

daytrip_long <- daytrip_long %>%
  rename(long_trip_count = n)

daytrip_short <- daytrip_short %>%
  rename(short_trip_count = n)

daytripper <- left_join(daytrip_long, daytrip_short, key = "day_of_week")





```


