---
title: "Lab5Individual"
author: "Rachael Robinson"
date: "2/13/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::kable
# Load tidyverse
library(tidyverse)
library(tidyverse)
library(knitr)
# Load and view the data
milk <- read.csv('state_milk_production.csv')


milk <- milk %>%
  mutate(milk_million = milk_produced/1000000)

milk1998 <- milk %>%
  filter(year == 1998)


```

```{r, echo = FALSE, warning = FALSE}
ggplot(data = milk1998, mapping = aes(x=milk_million)) +
  geom_density(bins=11, size = 1.5, color = "blue") +
  ggtitle("Distribution of US milk production in 1970") +
  theme_minimal() +
  xlab("Milk in millions of pounds")
```

```{r, include = FALSE}
  meanmilk <- mean(milk1998$milk_million)
  medmilk <- median(milk1998$milk_million)
  meanmilk
  medmilk
```

The plot shows the distribution of milk production in the US for the year 1998. From the plot, it's clear that the density is higher in the lower milk productions, from 0 to 4,000 million pounds. However, it appears there are a few states acting as outliers in the higher milk productions.

When evaluating the data for 1998, I have calcuated the average milk produced across all states to be 3145.22 millions of pounds.

The median was calculated to be 1411.5 millions of pounds.

The large discrepency between the median and the mean, along with the outliers observed, indicates that the mean is skewed upward.

The following is a table representing the top 5 states for milk production in the year 1998:

```{r, results = 'asis', echo = FALSE}

mostmilk <- milk1998 %>%
  select(state, milk_million) %>%
  arrange(desc(milk_million))

kable(mostmilk[1:5,])
```

This indicates that the highest producing state in 1998 was California with 27620 million pounds of milk.

This table shows the 5 states with the lowest milk production in the year 1998:

```{r, results = 'asis', echo = FALSE}

mostmilk <- milk1998 %>%
  select(state, milk_million) %>%
  arrange(desc(milk_million))

kable(mostmilk[45:50,])

```

This indicates that the lowest producing state in 1998 was Alaska with 14 million pounds of milk.
